ODGI: understanding pangenome graphs

### 2021-10-12 manuscript/main.tex
odgi viz/draw

By visualizing pangenome graphs, we can gain insight into the relationship between genomes.
odgi viz applies binning and direct rendering to a raster image to generate a 1D representation able to scale up to gigabase pangenome graphs (Figure 1.A).
However, pangenome graphs may present complex structures to visualize: odgi sort finds the best node order to simplify 1D graph visualization by applying several cutting-edge sorting algorithms.
odgi draw extends the visualization in 2D, taking in input the layouts built by odgi layout.

@@

% viz / draw
These provide scalable ways of generating human-viewable pictures of the high-level structure of the graph.
% interesting that we don't go base-level at all... we use vg for that, or STM

>>>>

@@

% interesting that we don't go base-level at all... we use vg for that, or STM

Pangenome graph visualization is one of the first steps to gain insight into the mutual relationship between the sequences and their variation.
\texttt{odgi viz} and \texttt{odgi draw} provide scalable ways of generating human-viewable pictures of the high-level structure of the graph.

\texttt{odgi viz} offers a binned, linearized rendering in 1 dimension (1D).

% Explain odgi viz's visualization

However, complex, nonlinear graph structures are difficult to display and interpret in a convenient low number of dimensions.
To overcome such a limitation, \texttt{odgi viz} supports multiple visualization modes, making it easy to grasp the properties of the graph and its complexity.

% Explain odgi viz's color by path position
% Explain odgi viz's color by strandness
% Explain odgi viz's color by coverage

\texttt{odgi draw} extends the visualization in 2 dimensions (2D), taking in input the layouts built by \texttt{odgi layout}.

% Explain odgi draw's visualization
